### 2021-10-22 reviews/01_ELE/01_draft_revised.tex
diff files

Here, we present a general method to recommend possible trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. There is a plurality of
measures of species similarities that can be used for metaweb
reconstruction (see \emph{e.g.} Morales-Castilla \emph{et al.} 2015);
however, phylogenetic proximity has several desirable properties when
working at large scales. Gerhold \emph{et al.} (2015) made the point
that phylogenetic signal captures diversification of characters (large
macro-evolutionary process), but not necessarily community assembly
(fine ecological process); Dormann \emph{et al.} (2010) previously found
very similar conclusions. Interactions tend reflect a phylogenetic
signal because they have a conserved pattern of evolutionary convergence
that encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares \emph{et al.} 2009; Mouquet \emph{et al.} 2012), and -
most importantly - retain this signal even when it is not detectable at
the community scale (Hutchinson \emph{et al.} 2017; Poisot \& Stouffer
2018). Finally, species interactions at macro-ecological scales seem to
respond mostly to macro-evolutionary processes (Price 2003); which is
evidenced by the presence of conserved backbones in food webs (Dalla
Riva \& Stouffer 2016), strong evolutionary signature on prey choice
(Stouffer \emph{et al.} 2012), and strong phylogenetic signature in food
web intervality (Eklöf \& Stouffer 2016). Phylogenetic reconstruction
has also previously been used within the context of ecological networks,
namely understanding ancestral plant-insect interactions (Braga \emph{et
al.} 2021). Taken together, these considerations suggest that
phylogenies can reliably be used to transfer knowledge on species
interactions.

@@

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (Maiorano \emph{et al.} 2020; known
interactions; \textbf{Maiorano2020TetEu?}) and, based on phylogenetic
relationships between mammals globally (\emph{i.e.}, phylogenetic tree
Upham \emph{et al.} 2019), infer a metaweb for the Canadian mammalian
species pool (interactions are treated as unknown in this instance).
Following the definition of Dunne (2006), a metaweb is a network
analogue to the concept of a regional species pool. Specifically, a
metaweb is an inventory of all \emph{possible} interactions within
species likely to occurr within a spatially delimited area (the network
\(\gamma\)-diversity, in a sense). The metaweb is, therefore, \emph{not}
a prediction of the food web at any specific locale within the frontiers
of the spatial area it recovers, and will in fact have a different
structure (notably by having a larger connectance; see \emph{e.g.}
\textbf{Wood2015EffSpa?}). These local food webs are expected to be a
subset of both the species and the interactions of their metaweb, and
have been called ``metaweb realizations'' (Poisot \emph{et al.} 2015).
The difference between a food web at a specific location and the metaweb
has to do with a variety of mechanisms, including species co-occurrence,
local environmental conditions, and local distribution of functional
traits. Nevertheless, the metaweb represents the total of functional,
phylogenetic, and macroecological processes (Morales-Castilla \emph{et
al.} 2015), and therefore still holds valuable ecological information.
Because the metaweb can be down-sampled given appropriate knowledge of
local species composition (the network \(\alpha\)-diversity, in a
sense), it is possible to infer what may drive the structure of food
webs at finer spatial scales. This has been done for example for
tree-gallers-parasitoid systems (\textbf{Gravel2018BriElt?}), fish
trophic interactions (Albouy \emph{et al.} 2019), tetrapods trophic
interactions (O'Connor \emph{et al.} 2020), and crop-pests networks
(Grünig \emph{et al.} 2020). Whereas the original metaweb definition was
based on presence/absence, we focus on \emph{probabilistic} metawebs;
not only does our method recommed interactions that may exist, it gives
each interaction a score that is mathematically equivalent to the chance
of success of a Bernoulli trial (see \emph{e.g.} Poisot \emph{et al.}
2016), which allows properly weigh interactions as a function of how
likely they are.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the result, when validated against
known but fracational data of trophic interactions between Canadian
mammals, our model achieves a predictive accuracy of approx. 91\%. It
should be reiterated that the framework presented in
fig.~\ref{fig:concept} is amenable to changes; notably, the measure of
similarity may not be phylogeny, and can be replaced by information on
foraging (Beckerman \emph{et al.} 2006), cell-level mechanisms
(Boeckaerts \emph{et al.} 2021), or a combination of traits and
phylogenetic structure (Stock 2021).

@@

We use data from the European metaweb assembled by
(\textbf{Maiorano2020TetEu?}).

@@

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the IUCN checklist, and selected the terrestrial and
semi-aquatic species (this corresponds to the same selection that was
applied by (\textbf{Maiorano2020TetEu?}) in the European metaweb). The
IUCN names were, as previously, reconciled against GBIF to have an exact
match to the taxonomy.

@@

The crux of the method is the transfer of knowledge of a known network,
in order to predict interactions between species from another location.
In fig.~\ref{fig:concept}, we give a high-level overview of the
approach; in the example around which this manuscript is built
(leveraging detailed knowledge about binary trophic interactions between
Mammalia in Europe to predict the less known trophic interactions
between closely phylogenetically related Mammalia in Canada), we use a
series of specific steps for network embedding, trait inference, network
prediction and thresholding.

@@

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson
\emph{et al.} 2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl}
(\textbf{Dansereau2021SimJl?}). The network embedding and analysis is
done using \texttt{EcologicalNetworks.jl} (Poisot \emph{et al.} 2019;
Banville \emph{et al.} 2021). The phylogenetic simulations are done
using \texttt{PhyloNetworks.jl} (Solís-Lemus \emph{et al.} 2017) and
\texttt{Phylo.jl} (Reeve \emph{et al.} 2016). A complete
\texttt{Project.toml} file specifying the full tree of dependencies is
available alongside the code. This material also includes a fully
annotated copy of the entire code required to run this project
(describing both the intent of the code and discussing some technical
implementation details), a vignette for every step of the process, and a
series of Jupyter notebooks with the text and code. The pipeline can be
executed on a laptop in a matter of minutes, and therefore does not
require extensive computational power.

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016). In addition, recent advances show that the
latent variables produced this way can be used to predict \emph{de novo}
network edges. Interstingly, the latent variables do not need to be
prouced by decomposing the network itself; in a recent contribution,
Runghen \emph{et al.} (2021) show that deep artificial neural networks
are able to reconstruct the left and right subspaces of a RDPG, in order
to predict human movement networks from individual/location metadata.
This is an exciting opportunity, as it opens up the possibility of using
additional predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom
\emph{et al.} 2021b) and to allow highly accurate prediction of the
interactions within a single network (Poisot \emph{et al.} 2021b). Under
SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{L}\mathbf{\Sigma}\mathbf{R'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{L}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{R}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{L}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{R}'\) a \(n \times r\) semi-unitary matrix.

@@

A RDPG estimates the probability of observing interactions between nodes
(species) as a function of the nodes' latent variables. The latent
variables used for the RDPG, called the left and right subspaces, are
defined as \(\mathscr{L} = \mathbf{L}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{R}\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R}' = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R}' \approx \mathbf{A}\). Using a rank of 1 for
the t-SVD provides a first-order approximation of the network.

@@

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
(\textbf{Maiorano2020TetEu?}) and O'Connor \emph{et al.} (2020), this
seems like a reasonable assumption, although one that does not hold for
all metawebs. We used the thresholding approach presented in Poisot
\emph{et al.} (2021b), and picked a cutoff that maximized Youden's \(J\)
statistic (a measure of the informedness (trust) of predictions; Youden
(1950)); the resulting cutoff was 0.22, and gave an accuracy above 0.99.

@@

@@

Second, we looked for all species in the Canadian pool known to the
Global Biotic Interactions (GLoBI) database (Poelen \emph{et al.} 2014),
and extracted their known interactions. Because GLOBI aggregates
observed interactions, it is not a \emph{networks} data source, and
therefore the only information we can reliably extract from it is that a
species pair \emph{was reported to interact at least once}. This last
statement should yet be taken with caution, as some sources in GLOBI
(\emph{e.g.} Thessen \& Parr 2014) are produced through text analysis,
and therefore may not document direct evidence of the interaction.
Nevertheless, should the predictive model work, we would expect that a
majority of interactions known to GLOBI would also be predicted. After
performing this check, we set the probability of all interactions known
to GLOBI (366 in total, 33 of which were not predicted by the model, for
a success rate of 91\%) to 1.

@@

Dallas \emph{et al.} (2017) suggest that most links in ecological
networks may be cryptic, \emph{i.e.} uncommon or otherwise hard to
observe. This argument essentially echoes Jordano (2016b): the sampling
of ecological interactions is difficult because it requires first the
joint observation of two species, and then the observation of their
interaction. In addition, it is generally expected that weak or rare
links would be more common in networks (Csermely 2004), compared to
strong, persistent links; this is notably the case in food chains,
wherein many weaker links are key to the stability of a system (Neutel
\emph{et al.} 2002). In the light of these observations, the results in
fig.~\ref{fig:inflation} are not particularly surprising: we expect to
see a surge in these low-probability interactions under a model that has
a good predictive accuracy. Because the predictions we generate are by
design probabilistic, then one can weigh these rare links appropriately.
In a sense, that most ecological interactions are elusive can call for a
slightly different approach to sampling: once the common interactions
are documented, the effort required in documenting each rare interaction
may increase exponentially. Recent proposals suggest that machine
learning algorithms, in these situations, can act as data generators
(Hoffmann \emph{et al.} 2019): in this perspective, high quality
observational data can be supplemented with synthetic data coming from
predictive models, which increases the volume of information available
for inference.

Cirtwill \emph{et al.} (2019) previously made the point that network
inference techniques based on Bayesian approaches would perform far
better in the presence of an interaction-level informative prior; the
desirable properties of such a prior would be that it is expressed as a
probability, preferably representing a Bernoulli event, the value of
which would be representative of relevant biological processes
(probability of predation in this case). We argue that the probability
returned at the very last step of our framework may serve as this
informative prior; indeed, the output of our analysis can be used in
subsequent steps, also possibly involving expert elicitation to validate
some of the most strongly recommended interactions. One important
\emph{caveat} to keep in mind when working with interaction inference is
that interactions can never really be true negatives (in the current
state of our methodological framework and data collection limitations);
this renders the task of validating a model through the usual
application of binary classification statistics very difficult (although
see Strydom \emph{et al.} 2021a for a discussion of alternative
suggestions). The other way through which our framework can be improved
is by substituting the predictors that are used for transfer. For
example, in the presence of information on species traits that are known
to be predictive of species interactions, one might want to rely on
functional rather than phylogenetic distances -- in food webs, body size
(and allometrically related variables) has been established as such a
variable (Brose \emph{et al.} 2006); the identification of relevant
functional traits is facilitated by recent methodological developments
(Rosado \emph{et al.} 2013). It should be noted that Xing \& Fayle
(2021) highlight phylogenetic relatedness as one of the core components
of network comparison at the global scale. In this case study, we have
embedded the original metaweb using t-SVD, because it lends itself to a
RDPG reconstruction, which is known to capture the consequences of
evolutionary processes (Dalla Riva \& Stouffer 2016); this being said,
there are others ways to embed graphs (Cai \emph{et al.} 2017; Arsov \&
Mirceva 2019; Cao \emph{et al.} 2019), which can be used as
alternatives.

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, our approach must
contend with two interesting problems. The first is the limit of the
metaweb to embed and transfer. If the initial metaweb is too narrow in
scope, notably from a taxonomic point of view, the chances of finding
another area with enough related species to make a reliable inference
decrease. This is notably true if the metaweb is assembled in an area
with mostly endemic species. Conversely, the metaweb should be reliably
filled, which assumes that the \(S^2\) interactions in a pool of \(S\)
species have been examined, either through literature surveys or expert
elicitation. The second problem is to determine which area should be
used to infer the new metaweb in, as this determines the species pool
that must be used. In our application, we focused on the mammals of
Canada. The upside of this approach is that information at the country
level is likely to be required by policy makers and stakeholders for
their biodiversity assessment, as each country tends to set goals at the
national level (Buxton \emph{et al.} 2021) for which quantitative
instruments are designed (Turak \emph{et al.} 2017), with specific
strategies often enacted at smaller scales (Ray \emph{et al.} 2021). Yet
these national divisions, in large parts of the world, reflect nothing
except for the legacy of settler colonialism; indeed, the use of
ecological data is not an apolitical act (Nost \& Goldstein 2021).
Operating under the framework of national boundaries, and relying on
data infrastructures designed to answer questions within them, must be
done under the clear realization that this logic contributed to the
ongoing biodiversity crisis (Adam 2014), can reinforce environmental
injustice (Choudry 2013; Domínguez \& Luoma 2020), and on Turtle Island
especially, will probably end up being replaced by Indigenous principles
of land management (Eichhorn \emph{et al.} 2019; No'kmaq \emph{et al.}
2021).

>>>>

Here, we present a general method to recommend possible trophic
interactions, relying on the transfer learning of network
representations, specifically by using similarities of species in a
biologically/ecologically relevant proxy space (\emph{e.g.} shared
morphology or ancestry). Transfer learning is a machine learning
methodology that uses the knowledge gained from solving one problem and
applying it to a related (destination) problem (Pan \& Yang 2010; Torrey
\& Shavlik 2010). In this instance, we solve the problem of predicting
trophic interactions between species, based on knowledge extracted from
another species pool for which interactions are known by using
phylogenetic structure as a medium for transfer. There is a plurality of
measures of species similarities that can be used for metaweb
reconstruction (see \emph{e.g.} Morales-Castilla \emph{et al.} 2015);
however, phylogenetic proximity has several desirable properties when
working at large scales. Gerhold \emph{et al.} (2015) made the point
that phylogenetic signal captures diversification of characters (large
macro-evolutionary process), but not necessarily community assembly
(fine ecological process); Dormann \emph{et al.} (2010) previously found
very similar conclusions. Interactions tend reflect a phylogenetic
signal because they have a conserved pattern of evolutionary convergence
that encompasses a wide range of ecological and evolutionary mechanisms
(Cavender-Bares \emph{et al.} 2009; Mouquet \emph{et al.} 2012), and -
most importantly - retain this signal even when it is not detectable at
the community scale (Hutchinson \emph{et al.} 2017; Poisot \& Stouffer
2018). Finally, species interactions at macro-ecological scales seem to
respond mostly to macro-evolutionary processes (Price 2003); which is
evidenced by the presence of conserved backbones in food webs (Dalla
Riva \& Stouffer 2016; Mora \emph{et al.} 2018), strong evolutionary
signature on prey choice (Stouffer \emph{et al.} 2012), and strong
phylogenetic signature in food web intervality (Eklöf \& Stouffer 2016).
Phylogenetic reconstruction has also previously been used within the
context of ecological networks, namely understanding ancestral
plant-insect interactions (Braga \emph{et al.} 2021). Taken together,
these considerations suggest that phylogenies can reliably be used to
transfer knowledge on species interactions.

@@

Our methodology is outlined in fig.~\ref{fig:concept}, where we provide
an illustration based on learning the embedding of a metaweb of trophic
interactions for European mammals (known interactions; Maiorano \emph{et
al.} 2020b, a) and, based on phylogenetic relationships between mammals
globally (\emph{i.e.}, phylogenetic tree Upham \emph{et al.} 2019),
infer a metaweb for the Canadian mammalian species pool (interactions
are treated as unknown in this instance). Following the definition of
Dunne (2006), a metaweb is a network analogue to the concept of a
regional species pool. Specifically, a metaweb is an inventory of all
\emph{possible} interactions within species likely to occurr within a
spatially delimited area (the network \(\gamma\)-diversity, in a sense).
The metaweb is, therefore, \emph{not} a prediction of the food web at
any specific locale within the frontiers of the spatial area it
recovers, and will in fact have a different structure (notably by having
a larger connectance; see \emph{e.g.} Wood \emph{et al.} 2015). These
local food webs are expected to be a subset of both the species and the
interactions of their metaweb, and have been called ``metaweb
realizations'' (Poisot \emph{et al.} 2015). The difference between a
food web at a specific location and the metaweb has to do with a variety
of mechanisms, including species co-occurrence, local environmental
conditions, and local distribution of functional traits. Nevertheless,
the metaweb represents the total of functional, phylogenetic, and
macroecological processes (Morales-Castilla \emph{et al.} 2015), and
therefore still holds valuable ecological information. Because the
metaweb can be down-sampled given appropriate knowledge of local species
composition (the network \(\alpha\)-diversity, in a sense), it is
possible to infer what may drive the structure of food webs at finer
spatial scales. This has been done for example for
tree-gallers-parasitoid systems (Gravel \emph{et al.} 2018), fish
trophic interactions (Albouy \emph{et al.} 2019), tetrapods trophic
interactions (O'Connor \emph{et al.} 2020), and crop-pests networks
(Grünig \emph{et al.} 2020). Whereas the original metaweb definition was
based on presence/absence, we focus on \emph{probabilistic} metawebs;
not only does our method recommed interactions that may exist, it gives
each interaction a score that is mathematically equivalent to the chance
of success of a Bernoulli trial (see \emph{e.g.} Poisot \emph{et al.}
2016), which allows properly weigh interactions as a function of how
likely they are.

Our case study shows that phylogenetic transfer learning is an effective
approach to the generation of probabilistic metawebs. This showcases
that although the components (species) that make up the Canadian and
European communities may be \emph{minimally} shared (the overall species
overlap is less than 4\%), if the medium (proxy space) selected in the
transfer step is biologically plausible, we can still effectively learn
from the known network and make biologically relevant predictions of
interactions. Indeed, as we detail in the result, when validated against
known but fracational data of trophic interactions between Canadian
mammals, our model achieves a predictive accuracy of approx. 91\%. It
should be reiterated that the framework presented in
fig.~\ref{fig:concept} is amenable to changes; notably, the measure of
similarity may not be phylogeny, and can be replaced by information on
foraging (Beckerman \emph{et al.} 2006), cell-level mechanisms
(Boeckaerts \emph{et al.} 2021), or a combination of traits and
phylogenetic structure (Stock 2021). Most importantly, although we focus
on a trophic system, it is an established fact that different types of
interactions do themselves interact (Kéfi \emph{et al.} 2012; see
\emph{e.g.} Kawatsu \emph{et al.} 2021). Future development of metaweb
inference techniques should cover the prediction of multiple interaction
types.

@@

We use data from the European metaweb assembled by Maiorano \emph{et
al.} (2020b).

@@

The destination problem to which we want to transfer knowledge is the
trophic interactions between mammals in Canada. We obtained the list of
extant species from the IUCN checklist, and selected the terrestrial and
semi-aquatic species (this corresponds to the same selection that was
applied by Maiorano \emph{et al.} (2020b) in the European metaweb). The
IUCN names were, as previously, reconciled against GBIF to have an exact
match to the taxonomy.

@@

The core point of our method is the transfer of knowledge of a known
ecological network, in order to predict interactions between species
from another location at which the network is unknown (or partially
known). In fig.~\ref{fig:concept}, we give a high-level overview of the
approach; in the example around which this manuscript is built
(leveraging detailed knowledge about binary trophic interactions between
Mammalia in Europe to predict the less known trophic interactions
between closely phylogenetically related Mammalia in Canada), we use a
series of specific steps for network embedding, trait inference, network
prediction and thresholding.

@@

The entire pipeline is implemented in \emph{Julia} 1.6 (Bezanson
\emph{et al.} 2017) and is available under the permissive MIT License at
\href{https://osf.io/2zwqm/}{\texttt{https://osf.io/2zwqm/}}. The
taxonomic cleanup steps are done using \texttt{GBIF.jl} (Dansereau \&
Poisot 2021). The network embedding and analysis is done using
\texttt{EcologicalNetworks.jl} (Poisot \emph{et al.} 2019; Banville
\emph{et al.} 2021). The phylogenetic simulations are done using
\texttt{PhyloNetworks.jl} (Solís-Lemus \emph{et al.} 2017) and
\texttt{Phylo.jl} (Reeve \emph{et al.} 2016). A complete
\texttt{Project.toml} file specifying the full tree of dependencies is
available alongside the code. This material also includes a fully
annotated copy of the entire code required to run this project
(describing both the intent of the code and discussing some technical
implementation details), a vignette for every step of the process, and a
series of Jupyter notebooks with the text and code. The pipeline can be
executed on a laptop in a matter of minutes, and therefore does not
require extensive computational power.

@@

The first step in transfer learning is to learn the structure of the
original dataset. In order to do so, we rely on an approach inspired
from representational learning, where we learn a \emph{representation}
of the metaweb (in the form of the latent subspaces), rather than a list
of interactions (species \emph{a} eats \emph{b}). This approach is
conceptually different from other metaweb-scale predictions (\emph{e.g.}
Albouy \emph{et al.} 2019), in that the metaweb representation is easily
transferable. Specifically, we use RDPG to create a number of latent
variables that can be combined into an approximation of the network
adjacency matrix. RDPG results are known to have strong phylogenetic
signal, and to capture the evolutionary backbone of food webs (Dalla
Riva \& Stouffer 2016); in other words, the latent variables of an RDPG
can be mapped onto a phylogenetic tree, and phylogenetically close
predators should share phylogenetically close preys. In addition, recent
advances show that the latent variables produced this way can be used to
predict \emph{de novo} network edges. Interstingly, the latent variables
do not need to be prouced by decomposing the network itself; in a recent
contribution, Runghen \emph{et al.} (2021) show that deep artificial
neural networks are able to reconstruct the left and right subspaces of
an RDPG, in order to predict human movement networks from
individual/location metadata. This is an exciting opportunity, as it
opens up the possibility of using additional predictors.

The latent variables are created by performing a truncated Singular
Value Decomposition (t-SVD) on the adjacency matrix. SVD is an
appropriate embedding of ecological networks, which has recently been
shown to both capture their complex, emerging properties (Strydom
\emph{et al.} 2021b) and to allow highly accurate prediction of the
interactions within a single network (Poisot \emph{et al.} 2021b). Under
SVD, an adjacency matrix \(\mathbf{A}\) (where
\(\mathbf{A}_{m,n}\in\mathbb{B}\) where 1 indicates predation and 0 an
absence thereof) is decomposed into three components resulting in
\(\mathbf{A} = \mathbf{U}\mathbf{\Sigma}\mathbf{V'}.\) Here,
\(\mathbf{\Sigma}\) is a \(m \times n\) diagonal matrix and contains
only singular (\(\sigma\)) values along its diagonal, \(\mathbf{U}\) is
a \(m \times m\) unitary matrix, and \(\mathbf{V}'\) a \(n \times n\)
unitary matrix. Truncating the SVD removes additional noise in the
dataset by omitting non-zero and/or smaller \(\sigma\) values from
\(\mathbf{\Sigma}\) using the rank of the matrix. Under a t-SVD
\(\mathbf{A}_{m,n}\) is decomposed so that \(\mathbf{\Sigma}\) is a
square \(r \times r\) diagonal matrix (whith \(1 \le r \le r_{full}\)
where \(r_{full}\) is the full rank of \(\mathbf{A}\) and \(r\) the rank
at which we truncate the matrix) containing only non-zero \(\sigma\)
values. Additionally, \(\mathbf{U}\) is now a \(m \times r\) semi
unitary matrix and \(\mathbf{V}'\) a \(n \times r\) semi-unitary matrix.
As an aside, most ecologists are indirectly familiar with SVD: Principal
Component Analysis is a special case of SVD, which is more sensitive to
numerical instabilities (see notably Shlens 2014).

@@

An RDPG estimates the probability of observing interactions between
nodes (species) as a function of the nodes' latent variables, and is a
way to turn a SVD (which decompose a matrix intro three) into two
matrices that can be multiplied to provide an approximation of the
network. The latent variables used for the RDPG, called the left and
right subspaces, are defined as
\(\mathscr{L} = \mathbf{U}\sqrt{\mathbf{\Sigma}}\), and
\(\mathscr{R} = \sqrt{\mathbf{\Sigma}}\mathbf{V}'\) -- using the full
rank of \(\mathbf{A}\), \(\mathscr{L}\mathscr{R} = \mathbf{A}\), and
using any smaller rank results in
\(\mathscr{L}\mathscr{R} \approx \mathbf{A}\). Using a rank of 1 for the
t-SVD provides a first-order approximation of the network. One advantage
of using a RDPG rather than a SVD is that the number of components to
estimate decreases; notably, one does not have to estimate the
eigenvalues of the SVD. Furthermore, the two subspaces can be directly
multiplied to yield a network.

@@

Because RDPG relies on matrix multiplication, the higher dimensions
essentially serve to make specific interactions converge towards 0 or 1;
therefore, for reasonably low ranks, there is no guarantee that the
values in the reconstructed network will be within the unit range. In
order to determine what constitutes an appropriate threshold for
probability, we performed the RDPG approach on the European metaweb, and
evaluated the probability threshold by treating this as a binary
classification problem, specifically assuming that both 0 and 1 in the
European metaweb are all true. Given the methodological details given in
Maiorano \emph{et al.} (2020b) and O'Connor \emph{et al.} (2020), this
seems like a reasonable assumption, although one that does not hold for
all metawebs. We used the thresholding approach presented in Poisot
\emph{et al.} (2021b), and picked a cutoff that maximized Youden's \(J\)
statistic (a measure of the informedness (trust) of predictions; Youden
(1950)); the resulting cutoff was 0.22, and gave an accuracy above 0.99.
In Supp. Mat. 1, we provide several lines of evidence that using the
entire network to estimate the threshold does not lead to overfitting;
that using a subset of species would yield the same threshold; that
decreasing the quality of the original data by adding of removing
interactions would minimally affect the predicive accuracy of RDPG
applied to the European metaweb; and that the networks reconstructed
from artifically modified data are reconstructured with the correct
ecological properties.

@@

It must be noted that despite bringing in a large amount of information
from the European species pool and interactions, the Canadian metaweb
has distinct structural properties. Following an approach similar to
Vermaat \emph{et al.} (2009), we show in Supp. Mat. 3 that not only can
we observed differences in a multivariate space between the European and
Canadian metaweb, we can also observe differences in the same space
between random subgraphs from these networks. These results line up with
the studies spatializing metawebs that have been discussed in the
introduction: changes in the species pool are driving local structural
changes in the networks.

@@

Second, we looked for all species in the Canadian pool known to the
Global Biotic Interactions (GLoBI) database (Poelen \emph{et al.} 2014),
and extracted their known interactions. Because GLOBI aggregates
observed interactions, it is not a \emph{networks} data source, and
therefore the only information we can reliably extract from it is that a
species pair \emph{was reported to interact at least once}. This last
statement should yet be taken with caution, as some sources in GLOBI
(\emph{e.g.} Thessen \& Parr 2014) are produced through text analysis,
and therefore may not document direct evidence of the interaction.
Nevertheless, should the predictive model work, we would expect that a
majority of interactions known to GLOBI would also be predicted. We
retrieved 366 interactions between mammals from the Canadian species
pool from GLOBI, 33 of which were not predicted by the model; this
results in a success rate of 91\%. After performing this check, we set
the probability of all interactions known to GLOBI to 1.

@@

Dallas \emph{et al.} (2017) suggest that most links in ecological
networks may be cryptic, \emph{i.e.} uncommon or otherwise hard to
observe. This argument essentially echoes Jordano (2016b): the sampling
of ecological interactions is difficult because it requires first the
joint observation of two species, and then the observation of their
interaction. In addition, it is generally expected that weak or rare
links would be more common in networks (Csermely 2004), compared to
strong, persistent links; this is notably the case in food chains,
wherein many weaker links are key to the stability of a system (Neutel
\emph{et al.} 2002). In the light of these observations, the results in
fig.~\ref{fig:inflation} are not particularly surprising: we expect to
see a surge in these low-probability interactions under a model that has
a good predictive accuracy. Because the predictions we generate are by
design probabilistic, then one can weigh these rare links appropriately.
In a sense, that most ecological interactions are elusive can call for a
slightly different approach to sampling: once the common interactions
are documented, the effort required in documenting each rare interaction
may increase exponentially. Recent proposals suggest that machine
learning algorithms, in these situations, can act as data generators
(Hoffmann \emph{et al.} 2019): in this perspective, high quality
observational data can be supplemented with synthetic data coming from
predictive models, which increases the volume of information available
for inference. Indeed, Strydom \emph{et al.} (2021a) suggest that
knowing the metaweb may render the prediction of local networks easier,
because it fixes an ``upper bound'' on which interactions can exist;
indeed, with a probabilistic metaweb, we can consider that the metaweb
represents an aggregation of informative priors on the interactions.

Related to the last point, Cirtwill \emph{et al.} (2019) show that
network inference techniques based on Bayesian approaches would perform
far better in the presence of an interaction-level informative prior;
the desirable properties of such a prior would be that it is expressed
as a probability, preferably representing a Bernoulli event, the value
of which would be representative of relevant biological processes
(probability of predation in this case). We argue that the probability
returned at the very last step of our framework may serve as this
informative prior; indeed, the output of our analysis can be used in
subsequent steps, also possibly involving expert elicitation to validate
some of the most strongly recommended interactions. One important
\emph{caveat} to keep in mind when working with interaction inference is
that interactions can never really be true negatives (in the current
state of our methodological framework and data collection limitations);
this renders the task of validating a model through the usual
application of binary classification statistics very difficult (although
see Strydom \emph{et al.} 2021a for a discussion of alternative
suggestions). The other way through which our framework can be improved
is by substituting the predictors that are used for transfer. For
example, in the presence of information on species traits that are known
to be predictive of species interactions, one might want to rely on
functional rather than phylogenetic distances -- in food webs, body size
(and allometrically related variables) has been established as such a
variable (Brose \emph{et al.} 2006); the identification of relevant
functional traits is facilitated by recent methodological developments
(Rosado \emph{et al.} 2013). It should be noted that Xing \& Fayle
(2021) highlight phylogenetic relatedness as one of the core components
of network comparison at the global scale. In this case study, we have
embedded the original metaweb using t-SVD, because it lends itself to an
RDPG reconstruction, which is known to capture the consequences of
evolutionary processes (Dalla Riva \& Stouffer 2016); this being said,
there are others ways to embed graphs (Cai \emph{et al.} 2017; Arsov \&
Mirceva 2019; Cao \emph{et al.} 2019), which can be used as
alternatives.

As Herbert (1965) rightfully pointed out, ``{[}y{]}ou can't draw neat
lines around planet-wide problems''; in this regard, our approach (and
indeed, any inference of a metaweb at large scales) must contend with
several interesting and interwoven families of problems. The first is
the limit of the metaweb to embed and transfer. If the initial metaweb
is too narrow in scope, notably from a taxonomic point of view, the
chances of finding another area with enough related species to make a
reliable inference decreases; this would likely be indicated by large
confidence intervals during ancestral character estimation, but the lack
of well documented metawebs is currently preventing from developing more
concrete guidelines. The question of phylogenetic relatedness and
dispersal is notably true if the metaweb is assembled in an area with
mostly endemic species, and as with every predictive algorithm, there is
room for the application of our best ecological judgement. Conversely,
the metaweb should be reliably filled, which assumes that the \(S^2\)
interactions in a pool of \(S\) species have been examined, either
through literature surveys or expert elicitation. Supp. Mat. 1 provides
some guidance as to the type of sampling effort that should be
prioritized. Although RDPG was able to maintain very high predictive
power when interactions where missing, the addition of false positive
interactions was immediately detected; this suggests that it may be
appropriate to err on the side of ``too many'' interactions when
constructing the initial metaweb to be transfered. The second series of
problems are related to determining which area should be used to infer
the new metaweb in, as this determines the species pool that must be
used. In our application, we focused on the mammals of Canada. The
upside of this approach is that information at the country level is
likely to be required by policy makers and stakeholders for their
biodiversity assessment, as each country tends to set goals at the
national level (Buxton \emph{et al.} 2021) for which quantitative
instruments are designed (Turak \emph{et al.} 2017), with specific
strategies often enacted at smaller scales (Ray \emph{et al.} 2021). And
yet, we do not really have a satisfying answer to the question of
``where does a food web stops?''; the current most satisfying solutions
involve examining the spatial consistency of network area relationships
(see \emph{e.g.} Galiana \emph{et al.} 2018; Galiana \emph{et al.} 2019,
2021), which is of course impossible in the absence of enough
information about the network itse;f.~This suggests that an \emph{a
